### 2021-08-05 Analyses/initial_exploration.Rmd
Manuscript w references + Data

---
title: "Initial Exploration"
author: "Nicolas Restrepo"
date: "9/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Variables of interest

This is an initial exploration of the student data. The dataset has four waves, and several questions that might be of interest for the question at hand. Here, I want to check what variables are available across waves and which do not have too many missing values. Let's load in the necessary packages:

```{r}
library(tidyverse)
library(naniar)
library(haven)
library(igraph)
library(lme4)
library(brms)
```

Now, let's try to read in the data:

```{r}
wave1 <- read_sav("~/Documents/global_local_influence/Data/PupilsWaveV.sav")
wave2 <- read_sav("~/Documents/global_local_influence/Data/PupilsWaveW_geanonimiseerd.sav")
wave3 <- read_sav("~/Documents/global_local_influence/Data/PupilsWaveX.sav")
wave4 <- read_sav("~/Documents/global_local_influence/Data/PupilsWaveY.sav")
```

Here, the variables are named a bit differently depending on the wave. Most variables in wave 1 end with an "a", while most in wave 2 end with a "b", and so the pattern continues. This is important for later joining the dataframes.

I am going to begin by narrowing down the dataframes into three main sections:

1) One related to alcohol, smoking, and drugs. 
2) Another related to hobbies and music. 
3) The last one involving the trust and altruism questions.

I am going to look for those questions in all four waves and see what the patterns of missingness look like.

```{r}
# First create a variable for id 
# Then filter for variables of interest
df1 <- wave1 %>% 
  mutate(id = paste0(schoolnr,namenr)) %>% 
  select(id, actdrug, actalcoh, actsmoke, contains('hobby'), contains('music'), contains('trust'))

glimpse(df1)
# Awesome, it looks like we expected.

# Same for wave2 
df2 <- wave2 %>% 
  mutate(id = paste0(schoolnr,namenr)) %>% 
  select(id, actdrugb, actalcob, actsmokb, contains('hobby'), contains('music'), contains('trust'))

glimpse(df2)

# Now for wave 3 
df3 <- wave3 %>% 
  mutate(id = paste0(schoolnr,namenr)) %>% 
  select(id, actdrugc, actalcoc, actsmokc, contains('hobby'), contains('music'), contains('trust'))

glimpse(df3)

# And last for wave 4 
df4 <- wave4 %>% 
  mutate(id = paste0(schoolnr,namenr)) %>% 
  select(id, actdrugd, actalcod, actsmokd, contains('hobby'), contains('music'), contains('trust'))

glimpse(df4)
```

Okay, now let's see how the missing values look like for each of the dataframes.

```{r}
map(list(df1, df2, df3, df4), vis_miss)
```

I think the best options right now are to look at the deviant behaviors, the first question of music, and the trust questions. I am sure I am going to find some useful ways of approaching these data. I am interested in the drop of missing values between wave 2 and 3. Did they start coding missing values any differently? These are questions that I will have to address. For now, let me think about how to build the networks.

## Building the networks

Constructing the networks here will be a bit tricky. There are many schools and many classes inside each school; in total I have 126 classes. The networks themselves are not very big. Thus, it'll be interesting to see whether I can pick up on some of the dynamics I am interested in given that most of the networks I will be able to construct are fundamentally 'local' in nature.

Let's look at just one network. I am going to focus here on friendship ties. There are multiple questions that get at this, so I will combine them.

```{r}
netdf <- wave1 %>% 
  filter(schoolnr == wave1$schoolnr[[1]]) %>% 
  select(namenr, schoolnr, contains('emosup'), contains('person'), 
         contains('friend'))

glimpse(netdf)
```
We only have 18 kids here. I just wonder how well we are going to capture "global" dynamics with this. Let me see if I can build the network.

```{r}
net_mat <- matrix(0, 18, 18)

for (j in 1:nrow(netdf)) {
  vect <- netdf[j,-c(1,2)] %>% as.numeric(.)
  matches <- vect[which(vect>0)]
  if (is_empty(matches)) { 
    net_mat[j,] <- 0
    } else {
    for (i in 1:length(matches)) {
      net_mat[j,matches[[i]]] <- 1
      }
  }
}

network <- graph_from_adjacency_matrix(net_mat, mode = 'directed')
plot(network)
```
Okay, this looks good. Now, the challenge is to be able to extend this to all schools. However, I am tempted to try to get multiwave data for one school and compute the averages. Having done that it might be easier to extend the analysis to encompass all schools.

Let's try to do it.

```{r}

get_scores <- function(x){
friends <- which(net_mat[x,]==1)
popular <- which.max(eigen_centrality(network)$vector)

avg_friends <- wave1 %>% 
  filter(schoolnr == wave1$schoolnr[[1]], 
         namenr %in% friends) %>% 
  select(actalcoh) %>% 
  summarise(avg = mean(actalcoh))

pop_score <- wave1 %>% 
  filter(schoolnr == wave1$schoolnr[[1]], 
         namenr %in% popular) %>% 
  select(actalcoh) %>% 
  summarise(pop_avg = mean(actalcoh))

return(c(popular_score = pop_score$pop_avg, 
         friends_score = avg_friends$avg))
}

map_df(c(1:18), get_scores)
```

A lot of missing values, even here. Can I put all this code together in a way that would allow me to iterate over schools?

```{r}
get_scores <- function(df, school,dv) {
    # Begin by taking the friendship columns 
    netdf <- df %>% 
    filter(schoolnr == school) %>% 
    select(namenr, schoolnr, contains('emosu'), contains('perso'), 
           contains('frien'))
  # Build the network 
  net_mat <- matrix(0, max(netdf$namenr), max(netdf$namenr))
  for (j in 1:max(netdf$namenr)) {
    vect <- netdf[netdf$namenr==j,-c(1,2)] %>% as.numeric(.)
    vect[which(vect > max(netdf$namenr))] <- NA
    matches <- unique(vect[which(vect>0)])

    if (is_empty(matches)) { 
      net_mat[j,] <- 0
    } else {
      for (i in 1:length(matches)) {
        net_mat[j,matches[[i]]] <- 1
      }
    }
  }
  network <- graph_from_adjacency_matrix(net_mat, mode = 'directed')
  # Function to go through all the kids in a school 
  calculate_school <- function(x){
    friends <- which(net_mat[x,]==1)
    cent_scores <- eigen_centrality(network)$vector
    popular <- which(cent_scores >= quantile(cent_scores, 0.85))

    avg_friends <- df %>% 
      filter(schoolnr == school, 
             namenr %in% friends) %>% 
      select(dv) %>% 
      rename(value = dv) %>% 
      summarise(avg = mean(value))

    pop_score <- df %>% 
      filter(schoolnr == school, 
             namenr %in% popular) %>% 
      select(dv) %>% 
      rename(value = dv) %>% 
      summarise(pop_avg = mean(value))

    return(c(namenr=x, 
             schoolnr = school, 
             popular_score = pop_score$pop_avg, 
             friends_score = avg_friends$avg))
  }
  pupil_names <- unique(netdf$namenr)
  scores_data <- map_df(pupil_names, calculate_school)

  return(scores_data)
}
list_schools <- unique(wave1$schoolnr)
```

I am ready to try to calculate the scores for all schools. Let's start small and move from there.

```{r}
ls_truncated <- list_schools[1:3]

trial_trunc <- map_df(ls_truncated, get_scores, df = wave1, dv = 'actalcoh')

glimpse(trial_trunc)
```

It works! Let's see what happens if we try to do all schools.

But before I am going to binarize the variable:

```{r}
wave1 <-  wave1 %>% 
  mutate(actalcoh = case_when(actalcoh == 1 ~ 0, 
                              actalcoh == 2 ~ 1, 
                              actalcoh == 3 ~ 1, 
                              actalcoh == 4 ~ 1, 
                              actalcoh == 5 ~ 1))

wave2 <-  wave2 %>% 
  mutate(actalcob = case_when(actalcob == 1 ~ 0, 
                              actalcob == 2 ~ 1, 
                              actalcob == 3 ~ 1, 
                              actalcob == 4 ~ 1, 
                              actalcob == 5 ~ 1))

wave3 <-  wave3 %>% 
  mutate(actalcoc = case_when(actalcoc == 1 ~ 0, 
                              actalcoc == 2 ~ 1, 
                              actalcoc == 3 ~ 1, 
                              actalcoc == 4 ~ 1, 
                              actalcoc == 5 ~ 1))

wave4 <-  wave4 %>% 
  mutate(actalcod = case_when(actalcod == 1 ~ 0, 
                              actalcod == 2 ~ 1, 
                              actalcod == 3 ~ 1, 
                              actalcod == 4 ~ 1, 
                              actalcod == 5 ~ 1))

```

```{r}
start <- Sys.time()
alc_scores_w1 <- map_df(list_schools, get_scores, df = wave1, dv = 'actalcoh')
end <- Sys.time()
```

We got all the schools. How much missing data is there here?

```{r}
vis_miss(alc_scores_w1)
```

A considerable amount in the friends' scores but nothing *too* bad. I should try to do it for the other waves and see how things look. Does the code behave well on the other waves?

```{r}
# Get list of schools for wave2 
list_schools_w2 <- unique(wave2$schoolnr)

# Annoyingly they are not the same
# And there is one value missing

list_schools_w2 <- list_schools_w2[-c(1)]
```

Before I can move ahead with this, I need to make the code more robust so that it takes different variables.

I've done that, let's see if it works now

```{r}
alc_scores_w2 <- map_df(list_schools_w2, get_scores, df = wave2, dv = 'actalcob')
```

```{r}
vis_miss(alc_scores_w2)
```
Shoot this looks very very good. Let me finish with the other 2.

```{r}
list_schools_w3 <- unique(wave3$schoolnr)
list_schools_w4 <- unique(wave4$schoolnr)
alc_scores_w3 <- map_df(list_schools_w3, get_scores, df = wave3, dv = 'actalcoc')
alc_scores_w4 <- map_df(list_schools_w4, get_scores, df = wave4, dv = 'actalcod')
```

Join the dataframes:

```{r}
# Join Wave 1 
df1 <- alc_scores_w1 %>% 
  mutate(namenr = as.double(namenr)) %>% 
  right_join(., wave1, by = c('namenr', 'schoolnr'))

# Join Wave 2 
df2 <- alc_scores_w2 %>% 
  mutate(namenr = as.double(namenr)) %>% 
  right_join(., wave2, by = c('namenr', 'schoolnr'))

# Join Wave 3 
df3 <- alc_scores_w3 %>% 
  mutate(namenr = as.double(namenr)) %>% 
  right_join(., wave3, by = c('namenr', 'schoolnr'))

# Join Wave 4 
df4 <- alc_scores_w4 %>% 
  mutate(namenr = as.double(namenr)) %>% 
  right_join(., wave4, by = c('namenr', 'schoolnr'))

```

This worked relatively well. Now, I need to find the IDs that occur in all waves. Remember that school number is different in Wave 2, so I need to find ways of dealing with that.

```{r}
# Create unique IDs
df1 <- df1 %>% mutate(id = paste0(schoolnr,namenr))
df2 <- df2 %>% filter(schoolnr != " ") %>% 
  mutate(schoolnr = str_trim(tolower(schoolnr))) %>% 
  mutate(id =paste0("0",schoolnr, namenr))
df3 <- df3 %>% mutate(id = paste0(schoolnr, namenr))
df4 <- df4 %>% mutate(id = paste0(schoolnr, namenr))

# Now let's find the intersections 
ids_all_waves <- Reduce(intersect, list(df1$id, df2$id, df3$id, df4$id))

alc_w1 <- df1 %>% 
  filter(id %in% ids_all_waves) %>% 
  select(id,friends_score, popular_score, actalcoh) %>% 
  rename(friends_w1 = friends_score, 
         popular_w1 = popular_score, 
         actalcoh_w1 = actalcoh)

alc_w2 <- df2 %>% 
  filter(id %in% ids_all_waves) %>% 
  select(id,friends_score, popular_score, actalcob) %>% 
  rename(friends_w2 = friends_score, 
         popular_w2 = popular_score, 
         actalcoh_w2 = actalcob)

alc_w3 <- df3 %>% 
  filter(id %in% ids_all_waves) %>% 
  select(id,friends_score, popular_score, actalcoc) %>% 
  rename(friends_w3 = friends_score, 
         popular_w3 = popular_score, 
         actalcoh_w3 = actalcoc)

alc_w4 <- df4 %>% 
  filter(id %in% ids_all_waves) %>% 
  select(id,friends_score, popular_score, actalcod) %>% 
  rename(friends_w4 = friends_score, 
         popular_w4 = popular_score, 
         actalcoh_w4 = actalcod)

df_wide <- left_join(alc_w1, alc_w2, by = "id") %>% 
  left_join(., alc_w3, by = "id") %>% 
  left_join(., alc_w4, by = "id")
```

Now, let me make the long dataframes.

```{r}
friends_long <- df_wide %>% 
  pivot_longer(cols = contains('friends'), names_to = "wave", values_to = "friends") %>% 
  select(id, friends)

popular_long <- df_wide %>% 
  pivot_longer(cols = contains('popular'), names_to = "wave", values_to = "popular") %>% 
  select(popular)

alc_long <- df_wide %>% 
  pivot_longer(cols = contains('actalcoh'), names_to = "wave", values_to = "actalcoh") %>% 
  select(actalcoh)

df_long <- cbind(friends_long, popular_long, alc_long)

df_long <- df_long %>% 
  mutate(popular = as.numeric(popular), 
         friends = as.numeric(friends), 
         id = as.factor(id))

dfl_clean <- drop_na(df_long) 
alcoh <- zap_labels(dfl_clean$actalcoh)
dfl_clean <- dfl_clean %>% 
  mutate(alcoh=zap_labels(alcoh))

write_csv(dfl_clean, "~/Documents/global_local_influence/Data/alcohol_long_binary.csv", col_names = TRUE)
```

```{r}
b1 <- glmer(alcoh ~ friends + popular + 1 + friends*popular + (1 | id), 
           data = dfl_clean, 
           family = "binomial")

summary(b1)
```